Implementing Real-time Transport Services over an Ossified Network

### 2016-05-22 paper.tex
Expand API discussion

In the IETF, the Transport Services (TAPS) working group is chartered to 
(1) develop a taxonomy of \emph{transport services}. That is, to identify the
features that comprise, and can be combined, to form complete transport
protocols; and (2) to develop an abstract API for applications to request
desirable services, allowing the system to select an appropriate transport
protocol based on application needs. It is hoped that this will loosen the
coupling between application and transport, so enabling deployment of new 
transport protocols.

@@

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point sending data
that cannot be used, because it depends on previous data that was lost. 
...complex trade-off, since you might want to send data that misses its 
   deadline, if it allows later data to be used.

...following from this, we have the concept of application-level framing
   to make the best use of data, irrespective of losses. This implies that
   the transport provides a \emph{message oriented} service that respects
   ADU boundaries.

...we note that many multimedia applications make use of multiple streams. 
   For example, audio and video are logically separate, etc.
   can be supported by opening multiple transport connections, but
   desirable to support a \emph{sub-stream} service to reduce overheads.

...finally, we note that \emph{congestion control} is important. 
   motivate: historically there has been a belief that real-time doesn't 
   do congestion control, needs an isochronous channel. 
   not practical for the Internet. 
   Also, experience has shown not necessary for many real-time applications.
   Yes, there are some that need constant bandwidth and cannot adapt, but
   many are adaptive. 
   Better user experience to adapt to fit the channel, rather than recover
   from loss after the channel discards packets that don't fit.

@@

Table \ref{tab:services} summarises the transport services required for
real-time multimedia applications. From this set of services, Table
\ref{tab:api} gives an outline API. As shown, the connection primitives
are the same as those of TCP. Hosts setup and teardown sockets using 
\texttt{create\_socket()} and \texttt{close\_socket()}. Servers
\texttt{bind()} to an address they wish to accept connections from, they
\texttt{listen()} on for incoming connections, and they \texttt{accept()}
an incoming connection. Clients \texttt{connect()} to a server.

The message-oriented service is exposed by the \texttt{send\_message()}
and \texttt{receive\_message()} functions. When sending messages,
applications specify a sequence number. Applications
set the receiver's play-out delay, using \texttt{set\_po\_delay()}, in
milliseconds. This is sent to the sender, for use as part of the deadline
estimation. The sender attaches a \texttt{lifetime} to each message,
which is combined with an estimate of the current round-trip-time, and the
time that the message has spent in the sending buffer, to determine if a
message will arrive within its lifetime. A message that won't arrive
within its lifetime is considered to have \emph{expired}. Applications
specify the application-level sequence number of a message they depend on
(\texttt{depends\_on}). If the message depends on a message that has
expired, then it too is considered to have expired. The partial
reliability service follows from the deadline and dependency services:
messages will be reliably transmitted until they expire.

@@

\begin{table}
  \centering
  \begin{tabular}{ll}
    \toprule
      Function & Parameters \\
    \midrule
      \texttt{create\_socket} &  \\
      \texttt{close\_socket} & \texttt{sd} -- Socket descriptor \\
    \midrule
      \texttt{bind} & \texttt{sd} -- Socket descriptor \\
                    & \texttt{addr} -- Address to bind to \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
      \texttt{listen} & \texttt{sd} -- Socket descriptor \\
            \texttt{accept} & \texttt{sd} -- Socket descriptor \\
                    & \texttt{addr} -- Address of peer \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
      \texttt{connect} & \texttt{addr} -- Address to connect to \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
    \midrule
      \texttt{set\_po\_delay} & \texttt{delay} -- Playout delay (in ms) \\
    \midrule
      \texttt{send\_message} & \texttt{sd} -- Socket descriptor \\
      & \texttt{buf} -- Message data \\
      & \texttt{len} -- Length of message data \\
      & \texttt{seq\_num} -- Sequence number \\
      & \texttt{lifetime} -- Lifetime of message (in ms) \\ 
      & \texttt{depends\_on} -- \texttt{seq\_num} of dependency \\
      & \texttt{substream} -- Substream identifier \\
      \texttt{receive\_message} & \texttt{sd} -- Socket descriptor \\
      & \texttt{buf} -- Buffer for message data \\
      & \texttt{len} -- Size of \texttt{buf} \\
      & \texttt{substream} -- Substream identifier \\
    \bottomrule
  \end{tabular}
  \caption{Outline transport API for real-time applications}
  \label{tab:api}
\end{table}

>>>>

In the IETF, the Transport Services (TAPS) working group is chartered to 
(1) develop a taxonomy of \emph{transport services}, that is, to identify the
features that comprise, and can be combined, to form complete transport
protocols; and (2) to develop an abstract API for applications to request
desirable services, allowing the system to select an appropriate transport
protocol based on application needs. It is hoped that this will loosen the
coupling between application and transport, so enabling deployment of new 
transport protocols.

%--------------------------------------------------------------------------------------------------
\subsection{Transport Services}

@@

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point sending data
that cannot be used, because it depends on previous data that was lost. 
\todo{...complex trade-off, since you might want to send data that misses its 
   deadline, if it allows later data to be used.}

\todo{...following from this, we have the concept of application-level framing
   to make the best use of data, irrespective of losses. This implies that
   the transport provides a \emph{message oriented} service that respects
   ADU boundaries.}

\todo{...we note that many multimedia applications make use of multiple streams. 
   For example, audio and video are logically separate, etc.
   can be supported by opening multiple transport connections, but
   desirable to support a \emph{sub-stream} service to reduce overheads.}

\todo{...finally, we note that \emph{congestion control} is important. 
   motivate: historically there has been a belief that real-time doesn't 
   do congestion control, needs an isochronous channel. 
   not practical for the Internet. 
   Also, experience has shown not necessary for many real-time applications.
   Yes, there are some that need constant bandwidth and cannot adapt, but
   many are adaptive. 
   Better user experience to adapt to fit the channel, rather than recover
   from loss after the channel discards packets that don't fit. }

@@

Table \ref{tab:services} summarises the transport services required for
real-time multimedia applications.

%--------------------------------------------------------------------------------------------------
\subsection{Abstract API}

Given the set of transport services outlined in Table \ref{tab:services}, 
we sketch an abstract API in  Table \ref{tab:api}. The primitives divide
into four categories:

\begin{itemize}
  \item Hosts setup and tear-down sockets using the \texttt{socket()}
    and \texttt{close()} functions, as in the standard Berkeley sockets
    API.

  \item The connection primitives are the same as those of TCP sockets.
    Servers \texttt{bind()} to a particular address and port, then
    \texttt{listen()} for and \texttt{accept()} incoming connections.
    Clients \texttt{connect()} to a server.

   \item Once the connection is established, the receiver application
     indicates the media play-out delay it is using, in milliseconds, via
     the \texttt{set\_po\_delay()} call. This specifies the time that the
     application will buffer data, to compensate for network timing jitter,
     before it is rendered to the user. The play-out delay is fed back to
     the sender host, for use as part of the media deadline estimation.

  \item Finally, message-oriented data transmission is exposed by the
    \texttt{send\_message()} and \texttt{recv\_message()} functions.  
    These expose a partially reliable message delivery service to the
    application, framing data such that either a complete message is
    delivered, or it is lost in its entirety.

    The \texttt{send\_message()} call takes four additional parameters
    compared to the standard Berkeley Sockets API. These are 1) a message
    sequence number, that can be used to re-order messages and detect 
    message loss; 2) a relative deadline, which is combined with an
    estimate of the current round-trip-time, and the time that the message
    has spent in the sending buffer, to determine if a message will arrive
    in time to be played-out; 3) the sequence number of any message on
    which this depends, for example the sequence number of a video I-frame
    on which a P-frame is predicted; and 4) a sub-stream identifier, used,
    for example, to differentiate audio, video, sub-title, control, and
    repair streams.

    The \texttt{recv\_message()} call returns the sequence number,
    dependency information, and sub-stream identifier along with any
    received message, allowing the receiver to direct it to the correct
    decoding queue.

    \todo{relate this to the transport services described earlier}

\end{itemize}

A message that won't arrive within its lifetime is considered to have
\emph{expired}. Applications specify the application-level sequence number
of a message they depend on (\texttt{depends\_on}). If the message depends
on a message that has expired, then it too is considered to have expired.
The partial reliability service follows from the deadline and dependency
services: messages will be reliably transmitted until they expire.

@@

\begin{table}
  \centering
  \begin{tabular}{ll}
    \toprule
      Function & Parameters \\
    \midrule
      \texttt{socket} & \texttt{af} -- Address family \\
                      & \texttt{st} -- Socket type \\
      \texttt{close} & \texttt{sd} -- Socket descriptor \\
    \midrule
      \texttt{bind} & \texttt{sd} -- Socket descriptor \\
                    & \texttt{addr} -- Address to bind to \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
      \texttt{listen} & \texttt{sd} -- Socket descriptor \\
            \texttt{accept} & \texttt{sd} -- Socket descriptor \\
                    & \texttt{addr} -- Address of peer \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
      \texttt{connect} & \texttt{addr} -- Address to connect to \\
                    & \texttt{addrlen} -- Length of \texttt{addr} \\
    \midrule
      \texttt{set\_po\_delay} & \texttt{delay} -- Playout delay (in ms) \\
    \midrule
      \texttt{send\_message} & \texttt{sd} -- Socket descriptor \\
      & \texttt{buf} -- Message data \\
      & \texttt{len} -- Length of message data \\
      & \texttt{seq\_num} -- Sequence number \\
      & \texttt{deadline} -- Relative deadline of message (in ms) \\ 
      & \texttt{depends\_on} -- \texttt{seq\_num} of dependency \\
      & \texttt{substream} -- Substream identifier \\
      \texttt{recv\_message} & \texttt{sd} -- Socket descriptor \\
      & \texttt{buf} -- Buffer for message data \\
      & \texttt{len} -- Size of \texttt{buf} \\
      & \texttt{substream} -- Substream identifier \\
      & \todo{return value, rather than param?}\\
    \bottomrule
  \end{tabular}
  \caption{Outline transport API for real-time applications 
          \todo{show return values of these calls}}
  \label{tab:api}
\end{table}
